DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and Procedures

### 2014-05-29 rules/rules14.tex
Updates to rules

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item May 25 - deadline for solver entrants
\item June 1 - deadline for final versions of solvers (but no new entrants after May 25).
\item July 17 - SMT workshop; completion of competition.
\item SMT-COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

@@

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.] We encourage submission of solvers and benchmarks for both tracks.
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

@@

It is important for competitors to track discussions on the SMT-COMP mailing
list, as clarifications and any updates to these rules will be posted there.

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted by its authors using the Star-Exec (\url{http://www.starexec.org}) service.  
The execution service enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a shared computer
cluster.  The execution service is provided free of
charge, but it does require registration to create a login account.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to the service.

For participation in SMT-COMP, a solver must be
uploaded to Star-Exec, made publicly available, and the organizers informed of its presence {\em and the logic divisions in which it is participating}.
A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

@@

@@

\header{Input and Output.}
%
Participating solvers must read a single benchmark script presented as its first argument.
Note that previous competitions presented benchmarks to a solver on its standard input
channel, and that the application track will send command inputs to the solver's standard input.
Accordingly, solvers are advised to support both input mechanisms. The organizers have easily adapted solvers that use only the standard input to Star-Exec with a simple wrapper script.

@@

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
\item In the main track, there may be a (single) \akey{set-option :print-success ...} command. Note that {\tt success} outputs are ignored by the post-processor used by the competition.
\item Scripts for the application track will have an initial $\akey{set-option :print-success true}$ command.
\item The script next optionally contains (any number of) $\akey{set-info}$ commands, which may be ignored by the solver.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands 
  and zero or more \akey{push 1} commands, 
  and followed by zero or more \akey{pop 1} commands.

%%%%%%%%%%%%%%%%%%% TODO: Need to define the set-info command.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.

@@

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
An abort scores the same as if the output is ``\texttt{unknown}''. 
Also, as a further measure to prevent misjudgments of solvers,
any  ``\texttt{success}'' outputs will be 
ignored.\footnote{
Note that SMT-LIBv2 requires that a solver produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false; ignoring the
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of previous competitions to run without changes.}

@@

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction, as was done in 2012. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviours of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehaviour, etc. that may happen on the model checker side.
\end{itemize}

\header{Input and Output.}
Participating solvers will be connected to a trace executor 
which will incrementally send commands to the standard input channel of the solver
and read responses from the standard output channel of the solver.
The commands will be taken from an \textbf{incremental benchmark script},
which is an SMT-LIB 2.0 script which satisfies the rules for an application script given above.
Note also that the trace executor will send a single 
\akey{set-option :print-success true} command to the solver before 
sending commands from the incremental benchmark script.

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There should be no \akey{get-unsat-core}, \akey{get-proof}, \akey{get-model}, \akey{get-assignments}, or \akey{get-values}commands in these benchmarks.

@@

\header{Main track - sequential mode.}
For the main track, the score on a given benchmark is the single $\langle e,n,m\rangle$ triple, with $m$ being the CPU time it takes to process the benchmark, that is, until the solver process terminates, and
\begin{itemize}
\item $e=0, n=1$ if the result of the check-sat command is correct,
\item $e=0, n=0$ if the result is {\tt unknown}, and
\item $e=1, n=0$ if the result is incorrect.
\end{itemize}

\header{Main track - parallel mode.}
The Star-Exec infrastructure reports both wall-clock time and CPU time for a given benchmark-solver job-pair. Each Star-Exec node has multiple (currently 4) cores; a multi-threaded solver can in principle make use of these cores to reduce its wall-clock time. Since the competition has not to date emphasized parallel solvers, we will use total CPU time (across all cores and subprocesses) as the measure of performance in the sequential mode of the main track; the imposed time out will also be on the CPU time.

@@

@@

@@

We define a competition-wide metric M as follows. Each solver competing in a division receives a number of points for its performance in that division: $(n/N)^2$, where $n$ is the number of benchmarks solved correctly and $N$ is the number of benchmarks used in that division. A solver's score for the whole competition is the sum of these values over all the 
divisions in which the solver participates.

The Olympic Games medals will be awarded as follows.
\begin{itemize}
\item The gold medal we will be awarded to the solver with the best competition-wide value of M.
\item The silver medal will be awarded to the solver that wins the (sequential) QF\_BV division according to the usual $\langle e',n',m'\rangle$ metric of subsection \ref{scoring}.
\item The bronze medal will be awarded to the solver with the second-best  competition-wide value of M.
\end{itemize}

{\em Rationale.} The metric M purposely emphasizes breadth of solver participation - a solver participating in many logics need not be the best in any one of them. The use of the square in the metric limits that somewhat - a solver still needs to do reasonably well compared to winners to be able to catch up by breadth of participation. The non-linear metric also gives added weight to completing close to all benchmarks in a division. The metric is also quite simple and the metric for a solver is independent of the performance of other solvers.

>>>>

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item June 1 - deadline for solver entrants
\item June 15 - deadline for final versions of solvers (but no new entrants after May 25).
\item July 17 - SMT workshop; completion of competition.
\item SMT-COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

@@

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.] We encourage submission of solvers and benchmarks for both tracks.
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

@@

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted by its authors using the StarExec (\url{http://www.starexec.org}) service.  
The execution service enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a shared computer
cluster.  The execution service is provided free of
charge, but it does require registration to create a login account.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to the service.

For participation in SMT-COMP, a solver must be
uploaded to StarExec, made publicly available, and the organizers informed of its presence {\em and the logic divisions in which it is participating}. For clarity, the solver should have just one configuration, or at least, that the default configuration is the one to be used for competition.
A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

@@

Competition solvers should be submitted with just one StarExec configuration so that it is
clear what configuration should be used for the competition. Separately named StarExec solvers can wrap the same underlying tool with a different configuration, if needed, for the application track.

@@

\header{Input and Output.}
%
A participating solver must read a single benchmark script presented as its first argument.
Note that previous competitions presented benchmarks to a solver on its standard input
channel, and that the application track will send command inputs to the solver's standard input.
Accordingly, solvers are advised to support both input mechanisms. The organizers have easily adapted solvers that use only the standard input to StarExec with a simple wrapper script.

@@

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
\item In the main track, there may be a (single) \akey{set-option :print-success ...} command. Note that {\tt success} outputs are ignored by the post-processor used by the competition; in 
the application track, the :print-success option may not be disabled.
\item The application track executor will send an initial $\akey{set-option :print-success true}$ command to the solver. 
\item The script next optionally contains (any number of) $\akey{set-info}$ commands, which may be ignored by the solver. Benchmarks are expected to have the following \akey{set-info} commands: \akey{(set-info :smt-lib-version 2.0)}, \akey{(set-info :category ...)}, \akey{(set-info :source ...)} (indicating the authorship or tool that produced the benchmark), and \akey{(set-info :status ...)}.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands; 
  there may also be zero or more \akey{push 1} commands, 
  and zero or more \akey{pop 1} commands, consistent with the uses of those commands in
  the SMT-LIB standard. Each \akey{check-sat} command may be preceded by a \akey{set-info :status} command that indicates the expected result of the 
  \akey{check-sat} command.

\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item All sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item Application track scripts may have $\akey{define-sort}$ and $\akey{define-sort}$ commands. 
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.

@@

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.

The StarExec service also limits the memory consumption of the solver processes. We expect the limit per solver-benchmark job-pair to be quite high, on the order of 100GB. The values of both the timeout limit and the memory limit are available to a solver process through environment variables defined and set by the StarExec service (see the StarExec user guide for more information).

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
Also, as a further measure to prevent misjudgments of solvers,
any  ``\texttt{success}'' outputs will be 
ignored.\footnote{
Note that SMT-LIBv2 requires that a solver produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false; ignoring the
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of previous competitions to run without changes.}

@@

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction, as was done in 2012. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehavior, etc. that may happen on the model checker side.
\end{itemize}
Note that the executor terminates processing the script upon receiving an incorrect response.

\header{Input and Output.}
Participating solvers will be connected to a trace executor 
which will incrementally send commands to the standard input channel of the solver
and read responses from the standard output channel of the solver.
The commands will be taken from an \textbf{incremental benchmark script},
which is an SMT-LIBv2 script which satisfies the rules for an application script given above.
Note also that the trace executor will send a single 
\akey{set-option :print-success true} command to the solver before 
sending commands from the incremental benchmark script.

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{define-sort}, 
\akey{define-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There will be no \akey{get-unsat-core}, \akey{get-proof}, \akey{get-model}, \akey{get-assignments}, or \akey{get-values} commands in these benchmarks.

@@

\header{Main track - sequential mode.}
For the main track, the score on a given benchmark is the single $\langle e,n,m\rangle$ triple, with $m$ being the CPU time it takes to process the benchmark, that is, until the solver process terminates, and
\begin{itemize}
\item $e=0, n=1$ if the result of the check-sat command is correct,
\item $e=1, n=0$ if the result is incorrect.
\item $e=0, n=0$ if the result is {\tt unknown} or the solver times-out or otherwise aborts without a response, and
\end{itemize}

\header{Main track - parallel mode.}
The StarExec infrastructure reports both wall-clock time and CPU time for a given benchmark-solver job-pair. Each StarExec node has multiple (currently 4) cores; a multi-threaded solver can in principle make use of these cores to reduce its wall-clock time. Since the competition has not to date emphasized parallel solvers, we will use total CPU time (across all cores and subprocesses) as the measure of performance in the sequential mode of the main track; the imposed time out will also be on the CPU time.

@@

Parallel track scores are not directly comparable to the sequential track scores. However, we will use them to determine a parallel-track winner in each division.

@@

\subsection{Judging}

It may happen that faulty benchmarks are discovered in the course of the competition. The judges reserve the right, with careful deliberation, to remove a benchmark from the competition
results if it is determined that the benchmark is faulty (e.g., syntactically invalid in a way that affects some solvers but not others). Organizers on solver teams will be recused from such decisions. Solver teams may appeal to the organizers if they suspect that a benchmark is faulty or otherwise not within the rules.

@@

We define a competition-wide metric M as follows. Each solver competing in a division receives a number of points for its performance in that division: $\langle e,(n/N)^2 \rangle$, where $e$ is the number of wrong answers to benchmarks, $n$ is the number of benchmarks solved correctly, and $N$ is the number of benchmarks used in that division. A solver's score for the whole competition is the sum of these tuples over all the 
divisions in which the solver participates. The best value of M is the one with the least total value of $e$, secondly, the largest total value for the second component of the tuple, and thirdly, the least value of $m$.

The Olympic Games medals will be awarded as follows.
\begin{itemize}
\item The gold medal we will be awarded to the solver with the best competition-wide value of M.
\item The silver medal will be awarded to the solver with the second-best competition-wide value of M.
\item The bronze medal will be awarded to the solver that wins the (sequential) QF\_BV division according to the usual $\langle e,n,m\rangle$ metric of subsection \ref{scoring}. 
\item The winner of the gold medal is excluded from consideration for the silver medal; the winners of the gold and silver medals are excluded from consideration for the bronze medal.
\end{itemize}

{\em Rationale.} The metric M purposely emphasizes breadth of solver participation - a solver participating in many logics need not be the best in any one of them. The use of the square in the metric limits that somewhat - a solver still needs to do reasonably well compared to winners to be able to catch up by breadth of participation. The non-linear metric also gives added weight to completing close to all benchmarks in a division. The metric is also quite simple and the metric for a solver is independent of the performance of other solvers. Time is omitted from the metric both to simplify the metric and because it is only of third importance in the regular competition metric and is difficult to compare across divisions.
